---
title: "Semantic Librarian Tutorial"
output: 
  html_document:
    toc: true
    toc_depth: 3
pkgdown:
  as_is: true
#vignette: >
#  %\VignetteIndexEntry{Semantic_Librarian_Tutorial}
#  %\VignetteEngine{knitr::rmarkdown}
#  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

```{r setup, message=FALSE, warning=FALSE}
library(RsemanticLibrarian)
```

This tutorial describes how to use the Semantic Librarian functions. Presntly, this package uses a corpus of abstracts from the APA database. A small sample of the required data files are automatically loaded with the package. These include:

- `article_df` a data frame containing abstract information (100 abstracts)
- `article_vectors` a matrix containing the semantic vectors for each article. Each row contains the vector for each abstract. The rows for each abstract correspond to the row entries in `article_df`
- `author_list` a vector containing the first 100 author names
- `dictionary_words` a vector containing 100 words, much smaller than actual dictionary
- `WordVectors` a matrix containing the semantic vectors for each word in `dictionary_words`.

The entire data set can be downloaded from the github repository for the R Shiny Semantic Libararian app [https://github.com/CrumpLab/SemanticLibrarian](https://github.com/CrumpLab/SemanticLibrarian)

## Searching the Semantic Librarian

Create some search terms:

```{r}
search <- "president of psychology"
```

### Check the dictionary

The `get_search_terms()` function converts the words in the search string to lower case, removes special characters, and finds the words in the search string that are in the dictionary. Search terms must be in the dictionary, in order to use existing semantic vectors for searching the data base.

```{r}
search <- get_search_terms(search,dictionary_words)
```

### Compute similarity of search terms to articles

The `get-search-article_similarities()` function returns the cosine similarity between the search terms and all of the abstracts in the data set. `query_type` can be set to 1 for compound search, 2 for AND search, and 3 for OR search.

```{r}
article_sims <- get_search_article_similarities(search_terms = search,
                                                d_words = dictionary_words,
                                                w_vectors = WordVectors,
                                                a_vectors = article_vectors,
                                                a_df = article_df,
                                                query_type = 1)

knitr::kable(head(article_sims), format="html", escape=FALSE)
```

### Plot similar articles using multi-dimensional scaling

The `get_mds_article_fits()` returns the most similar articles, up to the number specified by `num_articles`. This function also conducts mult-dimensional scaling, to produce a 2-d renderable semantic space, with article positions in the space as x, y coordinates. Finally, the `num_clusters` argument conducts k-means cluster analysis. The resulting data frame can be used for plotting, or to produce a table of the top articles. 

```{r}
MDS <- get_mds_article_fits(num_articles = 25, 
                            num_clusters = 3,
                            year_range = c(1900,2000),
                            input_df = article_sims,
                            a_vectors=article_vectors
                            )
```

Here is a simple plot of the abstract space using ggplot.

```{r}
library(ggplot2)
ggplot(MDS, aes(x=X,y=Y,color = cluster))+
  geom_point()
```

Alternatively, plotly could be used to generate a plot. Here, hovering over the point displays the title of abstract.

```{r}
library(plotly)
ggp <- ggplot(MDS, aes(x= X, y= Y,
                       color=as.factor(cluster),
                       text=wrap_title))+
  geom_hline(yintercept=0, color="grey")+
  geom_vline(xintercept=0, color="grey")+
  geom_point(aes(size=Similarity), alpha=.75)+
  theme_void()+
  theme(legend.position = "none")

ax <- list(
  title = "",
  zeroline = TRUE,
  showline = FALSE,
  showticklabels = FALSE,
  showgrid = FALSE
)

p <- ggplotly(ggp, tooltip="text",
              source = "article_SS_plot",
              hoverinfo="text") %>%
  layout(xaxis = ax, yaxis = ax,showlegend = FALSE) %>%
  #style(hoverinfo = 'title') %>%
  config(displayModeBar = F) %>%
  layout(xaxis=list(fixedrange=TRUE)) %>%
  layout(yaxis=list(fixedrange=TRUE))

p$elementId <- NULL
p
```



